Computing Restricted Isometry Constants via Mixed-Integer Semidefinite Programming

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
adapted to submission

\documentclass[peerreviewca,onecolumn]{IEEEtran}

@@

\noindent
\begin{abstract}
  One of the fundamental tasks in compressed sensing is finding the
  sparsest solution to an underdetermined system of linear equations. It is
  well known that although this problem is \NP-hard, under certain
  conditions it can be solved by using a linear program which minimizes the
  1-norm. The restricted isometry property has been one of the key
  conditions in this context. However, computing the best constants for
  this condition is itself \NP-hard. In this paper we propose a
  mixed-integer semidefinite programming approach for computing these
  optimal constants. This also subsumes sparse principal component
  analysis. Computational results with this approach allow to evaluate
  earlier semidefinite relaxations and show that the quality that can be
  obtained in reasonable time is limited.
\end{abstract}

@@

\begin{definition}\label{ripdef}
  The \emph{restricted isometry property} (RIP) of order~$k$ holds for a
  matrix $A \in \R^{m \times n}$ with constant~$\delta \geq 0$, if
  \begin{equation}\label{rip}
    (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
  \end{equation}
  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$.
\end{definition}

@@

\begin{definition}\label{ricdef}
  For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
  \eqref{rip},
  \begin{equation}\label{ric}
    \begin{aligned}
      \delta_k \define \argmin_{\delta \geq 0} \big\{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
      & \text{ for all }x \in \Sigma_k \big\},
    \end{aligned}
  \end{equation}
  is called \emph{restricted isometry constant} (RIC) of order~$k$.
\end{definition}

@@

Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
in the following sense: if the matrix~$A$ satisfies the RIP of order~$k$
with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
$\beta_k$, respectively, then this implies a restricted isometry constant
of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
So $A$ has a small RIC if and only if it has a large lower and a small upper 
RIC. But since in~\eqref{rip} usually only one of the two inequalities will 
be sharp, the asymmetric version yields more information about~$A$.

@@

\noindent
In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed-integer semidefinite program we want to 
propose and will also be checked against in the section on numerical
results.

@@

  Thus, the optimal objective values of \eqref{Proj} and \eqref{MISDP}
  agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
  suffices to show that \eqref{Proj} has an optimal solution of rank one.

@@

\begin{IEEEbiographynophoto}{Marc E. Pfetsch}
  obtained a Diploma degree in mathematics from the University of
  Heidelberg, Germany, in 1997. He received a PhD degree in
  mathematics in 2002 and the Habilitation degree in 2008 from TU
  Berlin, Germany. From 2008 to 2012 he was a Full Professor for
  Mathematical Optimization at TU Braunschweig, Germany. Since 04/2012
  he has been a Full Professor for Discrete Optimization at TU
  Darmstadt, Germany. His research interests are mostly in discrete
  optimization, in particular symmetry in integer programs, compressed
  sensing, and algorithms for mixed integer programs.
\end{IEEEbiographynophoto}

>>>>

\documentclass[journal,onecolumn]{IEEEtran}

@@

\begin{abstract}
  \noindent
  One of the fundamental tasks in compressed sensing is finding the
  sparsest solution to an underdetermined system of linear equations. It is
  well known that although this problem is \NP-hard, under certain
  conditions it can be solved by using a linear program which minimizes the
  1-norm. The restricted isometry property has been one of the key
  conditions in this context. However, computing the best constants for
  this condition is itself \NP-hard. In this paper we propose a
  mixed-integer semidefinite programming approach for computing these
  optimal constants. This also subsumes sparse principal component
  analysis. Computational results with this approach allow to evaluate
  earlier semidefinite relaxations and show that the quality that can be
  obtained in reasonable time is limited.
\end{abstract}

@@

\begin{definition}\label{ripdef}
  The \emph{restricted isometry property} (RIP) of order~$k$ holds for a
  matrix $A \in \R^{m \times n}$ with constant~$\delta \geq 0$, if
  \begin{equation}\label{rip}
    (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
    \quad\text{for all }x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}.
  \end{equation}
\end{definition}

@@

\begin{definition}\label{ricdef}
  For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
  \eqref{rip},
  \begin{equation}\label{ric}
    \begin{aligned}
      \delta_k \define \argmin_{\delta \geq 0} \big\{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 
      \quad\text{ for all }x \in \Sigma_k \big\},
    \end{aligned}
  \end{equation}
  is called \emph{restricted isometry constant} (RIC) of order~$k$.
\end{definition}

@@

Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
in the following sense: if the matrix~$A$ satisfies the RIP of order~$k$
with RIC~$\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
$\beta_k$, respectively, then this implies a restricted isometry constant
of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
So $A$ has a small RIC if and only if it has a large lower and a small upper 
RIC. But since in~\eqref{rip} usually only one of the two inequalities will 
be sharp, the asymmetric version yields more information about~$A$.

@@

\noindent
In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant~$\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed-integer semidefinite program we want to 
propose and will also be checked against in the section on numerical
results.

@@

  Thus, the optimal objective values of \eqref{Proj} and \eqref{MISDP}
  agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
  suffices to show that~\eqref{Proj} has an optimal solution of rank one.

@@

\begin{IEEEbiographynophoto}{Marc E. Pfetsch}
  obtained a Diploma degree in mathematics from the University of
  Heidelberg, Germany, in 1997. He received a PhD degree in
  mathematics in 2002 and the Habilitation degree in 2008 from TU
  Berlin, Germany. From 2008 to 2012 he was a Full Professor for
  Mathematical Optimization at TU Braunschweig, Germany. Since 04/2012
  he has been a Full Professor for Discrete Optimization at TU
  Darmstadt, Germany. His research interests are mostly in discrete
  optimization, in particular symmetry in integer programs, compressed
  sensing, and algorithms for mixed integer programs.
\end{IEEEbiographynophoto}
\vfill
\mbox{}
